---
title: "ADA Project_Godambe 12.02.22"
author: "Maya S. Godambe"
output: pdf_document
always_allow_html: true
---
######################################################
# Project: Forzen Meals, Education and Income
# Author: Maya S. Godambe 
# Edit date: December 02,2022
# Data:((<https://wwwn.cdc.gov/Nchs/Nhanes/2017-2018/DBQ_J.htm>)), ((<https://wwwn.cdc.gov/Nchs/Nhanes/2017-2018/DEMO_J.htm>))
#########################################################

```{r setup, include=FALSE}
#Loading packages
pacman::p_load(stargazer, DiagrammeR, table1, car, sandwich, MASS, dplyr, lmtest, SMPracticals, ggplot2, writexl, broom, stats,tidyr, foreign)

#loading XPT files from NHANES website
demo <- read.xport("DEMO_J.XPT")
foodies <- read.xport("DBQ_J.XPT")

#Combining XPT files into dataset
nhanes<- cbind(demo,foodies)
```


```{r clean data, include = FALSE}
#Piping data and cleaning
nhanes2018 <- nhanes%>%
dplyr::select(RIAGENDR, RIDRETH3,DBD910,RIDAGEYR,INDFMIN2, DMDEDUC2)%>%
 mutate(RIAGENDR = recode_factor(.x = RIAGENDR,
                                '1' = ' Male',
                                '2' = ' Female',
                                '7' = NA_character_,
                                '9' = NA_character_))%>%
    rename(gender = RIAGENDR)%>%
  
  # Recoding race as a factor, and renaming as race
 mutate(RIDRETH3 = recode_factor(.x = RIDRETH3,
                                '1' = ' Hispanic',
                                '2' = ' Hispanic',
                                '3' = ' White',
                                '4' = ' Black',
                                '6' = ' Asian',
                                '7' = ' Other Race - Including Multi-Racial',
                                '.' = NA_character_))%>%
    rename(race= RIDRETH3)%>%
  
# Recoding as a factor and renaming as income
# How to distribute the income levels in Figure 2. ((<https://sgp.fas.org/crs/misc/R44705.pdf>))
   mutate(INDFMIN2 = recode_factor(.x = INDFMIN2,
                                '1' = ' $0 to $24,999', 
                                '2' = ' $0 to $24,999',
                                '3' = ' $0 to $24,999',
                                '4' = ' $0 to $24,999', 
                                '5' = ' $0 to $24,999', 
                                '6' = ' $25,000 to $54,999',
                                '7' = ' $25,000 to $54,999',
                                '8' = ' $25,000 to $54,999', 
                                '9' = ' $55,000 to $74,999',
                                '10' = ' $55,000 to $74,999', 
                                '12' = NA_character_,
                                '13' = NA_character_, 
                                '14' = ' $75,000 to $99,999', 
                                '15' = ' $100,000 and over', 
                                '77' = NA_character_, 
                                '99' = NA_character_)) %>%
    rename(income = INDFMIN2)%>%
  
# Recoding as a factor and labeling as education
   mutate(DMDEDUC2 = recode_factor(.x = DMDEDUC2,
                                '1' = ' Less than 9th grade',
                                '2' = ' 9-12th grade no diploma',
                                '3' = ' High school graduate/GED or equivalent',
                                '4' = ' Some college or AA degree',
                                '5' = ' College graduate or above',
                                '7' = NA_character_,
                                '9' = NA_character_,
                                '.' = NA_character_))%>%
  rename(education = DMDEDUC2) %>%


 # Renaming age (years) 
rename(age= RIDAGEYR) 

# Looking at the numbe of na's in data set after recoding
sum(is.na(nhanes2018))

```

```{r removing counts of frozen meals (DBD910) higher than 60}
# Removing count of frozen meals that are greater than 60. For health 
# and ethical concerns. I am not a doctor nor public health professional yet.

nhanes2018 <- nhanes2018%>%
mutate(DBD910 = na_if(x = DBD910, y = "0"))%>%
mutate(DBD910 = na_if(x = DBD910, y = "6666"))%>%
mutate(DBD910 = na_if(x = DBD910, y = "9999"))%>%
mutate(DBD910 = na_if(x = DBD910, y = "90"))%>%
mutate(DBD910 = na_if(x = DBD910, y = "80"))%>%
mutate(DBD910 = na_if(x = DBD910, y = "60"))%>%

# Renaming as frozen
rename(frozen = DBD910) 
```



```{r removing rows with NA for any variable}
#Removing NA rows
nhanes2018$frozen<- as.numeric(nhanes2018$frozen)

# omitting any observations with na's in the data set (do not do this in the futrue)
nhanes.clean <- na.omit(nhanes2018)

# Summarizing the counts of data
summary(nhanes.clean)

```
# Figure 1. Logic Model for data selection/exclusion

```{r Figure 1. Logic model for data selection/exclusion}
#Figure 1 Logic model for data selection/exclusion
grViz("digraph flowchart {

      tab1 [label = '@@1'] # starting
      tab2 [label = '@@2'] # exclusion 1
      tab3 [label = '@@3'] # exclusion 2
      tab1 -> tab2 -> tab3
}
      [1]: 'Records received from NHANES 2017-2018 with answers for demographics and nutrition questions n= 9,254'
      [2]: 'Excluding 44 individuals who have eaten 60 or more frozen meals/pizza in the past 30 days n = 9,210'
      [3]:'Excluding 7,524 individuals with missing data on variables used in the analysis n= 1,686'
      ")

```
#Table 1.

```{r table 1}
# Adding labels to the variable on the table
label(nhanes.clean$income) <- "Annual Household Income"
label(nhanes.clean$gender)<-"Gender"
label(nhanes.clean$race)<-"Race"
label(nhanes.clean$education)<-"Education"
label(nhanes.clean$age)<-"Age"
label(nhanes.clean $frozen) <- "Number of frozen meals/pizza in past 30 days"

# Using table one to visualize distribution of variables
table1(~ gender + frozen + age + race + education + income,overall="Total",
       footnote='SD = standard deviation',
       render.missing = NULL,
       data = nhanes.clean,
       caption = "Table 1.  (n = 1686; NHANES 2017-2018)")
```
#Checking assumptions

```{r assumptions}
# Check shape of distribution of counts of cases using density plot and histogram
# data is right skewed and observations are independent
ggplot(nhanes.clean, aes(frozen))+
    geom_histogram(breaks=c(0,2,4,6,8,9,10,12,14,16,18,20,22,24,26,28,30:90), 
                   color="black", fill="red")

#The Mean does not equal the variance thus, negative binomial will be used.
mean(nhanes.clean$frozen)
var(nhanes.clean$frozen)
```
The Mean does not equal the variance (overdispersion), thus negative binomial

# Visualization of Varibales
```{r ditribution of age and number of frozen meals eaten}
# Creating scatter plot to check the distribution of age by number of frozen meals
nhanes.clean%>%
ggplot(aes(x= frozen, y = age)) + 
      geom_point(alpha=1) +
    theme(legend.position="none",plot.title = element_text(size=10)) +
    labs(x ="Number of Frozen Meals/Pizza eaten in the past 30 days" , y = 
           "Age (years)",
       title = "Figure 2. Number of frozen meals by highest level of 
       education (n = 1686; NHANES 2017-2018)") 
```
```{r box plots of distribution of different variables with outcome}
# Using ggplot2 Figure 3. Boxplot (Race with Number of Forzen meals eaten in the past 30 days)
ggplot(nhanes.clean, aes(x=race, y=frozen, fill=race)) + 
    geom_boxplot(alpha=1) +
    theme(legend.position="none",plot.title = element_text(size=10)) +
    scale_fill_brewer(palette="OrRd") +
  labs(y ="Number of Frozen Meals/Pizza eaten in the past 30 days" , x = "Race",
       title = "Figure 3. Number of frozen meals by race n = 1686") +
  stat_boxplot(geom ='errorbar', width = 0.6) 
# Using ggplot2 Figure 4. Boxplot (Gender with Number of Forzen meals eaten in the past 
# 30 days)
ggplot(nhanes.clean, aes(x=gender, y=frozen, fill=gender)) + 
    geom_boxplot(alpha=1) +
    theme(legend.position="none",plot.title = element_text(size=10)) +
    scale_fill_brewer(palette="OrRd") +
labs(y ="Number of Frozen Meals/Pizza eaten in the past 30 days" , x = "Gender",
       title = "Figure 4. Number of frozen meals by gender n = 1686") +
    stat_boxplot(geom ='errorbar', width = 0.6) 

# Using ggplot2 Figure 5. Boxplot (Income with Number of Forzen meals eaten in the past 
# 30 days)
ggplot(nhanes.clean, aes(x=income, y=frozen, fill=income)) + 
    geom_boxplot(alpha=1) +
    theme(legend.position="none",plot.title = element_text(size=10)) +
    scale_fill_brewer(palette="OrRd") +
  coord_flip() +
   labs(y ="Number of Frozen Meals/Pizza eaten in the past 30 days" , x = 
          "Household income",
       title = "Figure . Number of frozen meals by household income n = 1686") +
    stat_boxplot(geom ='errorbar', width = 0.6) 

# Using ggplot2 Figure 6. Boxplot (Education with Number of Forzen meals eaten in the past
# 30 days)
ggplot(nhanes.clean, aes(x=education, y=frozen, fill=education)) + 
    geom_boxplot(alpha=1) +
    theme(legend.position="none",plot.title = element_text(size=10)) +
    scale_fill_brewer(palette="OrRd") +
  coord_flip() +
  labs(y ="Number of Frozen Meals/Pizza eaten in the past 30 days" , x = 
         "Highest Level of education",
       title = "Figure 6. Number of frozen meals by highest level of education 
       n = 1686") +
    stat_boxplot(geom ='errorbar', width = 0.6) 

```



# Comparing Poisson Model against Negative Binomial
```{r indivudal models without cofactors}
# Unadjusted Poisson and negative binomial model of income and number of frozen meals/pizza's eaten in the past 30 days
summary(model.P <- glm(frozen ~ income ,  family = poisson, data =  nhanes.clean)) 
summary(model.2 <- glm.nb(model.1 <- glm(frozen ~ income , data = nhanes.clean)))

 # Unadjusted Poisson and negative binomial model of highest level of education and number of frozen meals/pizza's eaten in the past 30 days
summary(model.E <- glm(frozen ~ education ,  family = poisson, data =  nhanes.clean)) 
summary(model.3 <- glm.nb(model.1 <- glm(frozen ~ education , data = nhanes.clean)))
```
```{r likelihood ratio tests}
#run lrtest to compare models
lrtest(model.P, model.2)
lrtest(model.E, model.3)
```
The logliklihood and the significant p-value suggest that the negative binomial model is better and that there is overdispersion. (Not statistically equivalent)

# Full Model and colinearity test
```{r full model}
# Full negative binomial model adjusting for age, gender and race 
# technically education can confound income and vice versa
summary(model.genderageraceedu <- glm.nb(model.genderagerace <- glm(frozen ~ 
            income + education + gender + age + race , data = nhanes.clean)))

# Checking for multicolinearity
vif(model.genderageraceedu)
#VIF are less than 10. No indication of colinearity from this method.
```

# Exponentiating Models
```{r summary for indivudal models}
# Printing non-adjusted models with 95% CI and Incident rates
tidy(model.2, conf.int=TRUE, exponentiate = TRUE)
tidy(model.3, conf.int=TRUE, exponentiate = TRUE)

# Printing full model adjusted for age race, gender and education
tidy(model.genderageraceedu, conf.int=TRUE, exponentiate = TRUE)


# Main finding: As education and income increases, number of frozen meals/ pizzas decreases.
```

# Code for presentable tables, thank you Josh L. 

```{r presentable table 2, results='asis'}
library(xtable)
full.summ = summary(model.genderageraceedu)
full.summ.coef = coef(full.summ)
full.summ.conf = confint(model.genderageraceedu)
full.summ.bind = cbind(full.summ.coef, full.summ.conf)
full.summ.rr = cbind(exp(full.summ.bind[,c(1, 5, 6)]), full.summ.bind[,4])
colnames(full.summ.rr) = c('IRR', '2.5%', '97.5%', 'p-value')
full.table = xtable(full.summ.rr)
print(full.table, type = 'latex')
```

```{r presentable table 3,results='asis'}
library(xtable)
full.summ.2 = summary(model.2)
full.summ.coef.2 = coef(full.summ.2)
full.summ.conf.2 = confint(model.2)
full.summ.bind.2 = cbind(full.summ.coef.2, full.summ.conf.2)
full.summ.rr.2 = cbind(exp(full.summ.bind.2[,c(1, 5, 6)]), full.summ.bind.2[,4])
colnames(full.summ.rr.2) = c('IRR', '2.5%', '97.5%', 'p-value')
full.table.2 = xtable(full.summ.rr.2)
print(full.table.2, type = 'latex')
```

```{r presentable table 4 ,results='asis'}
full.summ.3 = summary(model.3)
full.summ.coef.3 = coef(full.summ.3)
full.summ.conf.3 = confint(model.3)
full.summ.bind.3 = cbind(full.summ.coef.3, full.summ.conf.3)
full.summ.rr.3 = cbind(exp(full.summ.bind.3[,c(1, 5, 6)]), full.summ.bind.3[,4])
colnames(full.summ.rr.3) = c('IRR', '2.5%', '97.5%', 'p-value')
full.table.3 = xtable(full.summ.rr.3)
print(full.table.3, type = 'latex')
```

